Simple repression draft

### 2020-10-22 doc/resubmission/appendix_02_bursty.tex
final pass before resubmission

To clean up Equation~\ref{eq:si_master_deltat} we can send the first term on the
right hand side to the left, and divide both sides by $\Delta t$. Upon taking
the limit where $\Delta t \rightarrow 0$ we can write
\begin{equation}
{d \over dt}p(m, t) = (m + 1) \gamma p(m + 1, t)
- m \gamma p(m, t)
+ k_i \sum_{m'=0}^{m-1} G_{m-m'} p(m', t) 
- k_i \sum_{m'=m + 1}^{\infty} G_{m'-m} p(m, t).
\end{equation}
Furthermore, given that the timescale for this equation is set by the mRNA
degradation rate $\gamma$ we can divide both sides by this rate, obtaining
\begin{equation}
{d \over d\tau}p(m, \tau) = (m + 1) p(m + 1, \tau)
- m p(m, \tau)
+ \lambda \sum_{m'=0}^{m-1} G_{m-m'} p(m', \tau) 
- \lambda \sum_{m'=m + 1}^{\infty} G_{m'-m} p(m, \tau),
\label{eq:si_master_ode}
\end{equation}
where we defined $\tau \equiv t \times \gamma$, and $\lambda \equiv k_i/\gamma$.
The last term in Eq.~\ref{eq:si_master_ode} sums all burst sizes except for a 
burst of size zero. We can re-index the sum to include this term, obtaining
\begin{equation}
\lambda \sum_{m'=m + 1}^{\infty} G_{m'-m} p(m, \tau) = \lambda p(m, t) \left[
\underbrace{\sum_{m'={m}}^{\infty}G_{m'-m}}
_{\text{re-index sum to include burst size zero}} -
\underbrace{G_0}_{\text{subtract extra added term}}\right].
\end{equation}
Given the normalization constraint of the geometric distribution, adding the
probability of all possible burst sizes --including size zero since we
re-indexed the sum-- allows us to write
\begin{equation}
\sum_{m'=m}^{\infty}G_{m'-m} - G_0 = 1 - G_0.
\end{equation}
Substituting this into Eq.~\ref{eq:si_master_ode} results in
\begin{equation}
{d \over d\tau}p(m, \tau) = (m + 1) p(m + 1, \tau)
- m p(m, \tau)
+ \lambda \sum_{m'=0}^{m-1} G_{m-m'} p(m', \tau) 
- \lambda p(m, \tau) \left[ 1 - G_0 \right].
\label{eq:si_master_ode_2}
\end{equation}
To finally get at a more compact version of the equation notice that the third
term in Eq.~\ref{eq:si_master_ode_2} includes burst from size $m'-m = 1$ to size
$m' - m = m$. We can include the term $p(m, t) G_0$ in the sum which allows
bursts of size $m' - m = 0$. This results in our final form for the chemical
master equation
\begin{align}
{d \over d\tau}p(m, \tau) = 
(m + 1) p(m+1, \tau)
- m p(m, \tau) - 
\lambda p(m, \tau)
+ \lambda \sum_{m^\prime=0}^m G_{m-m^\prime} p(m^\prime, \tau).
\label{eq:si_master_unreg}
\end{align}

In order to solve Eq.~\ref{eq:si_master_unreg} we will use the generating
function method~\cite{vanKampen2007}. The probability generating function
is defined as
\begin{align}
F(z,t) = \sum_{m=0}^\infty z^m p(m,t),
\end{align}
where $z$ is just a dummy variable that will help us later on to obtain the
moments of the distribution. Let us now multiply both sides of
Eq.~\ref{eq:si_master_unreg} by $z^m$ and sum over all $m$
\begin{equation}
\sum_m z^m {d \over d\tau} p(m, \tau) = 
\sum_m z^m \left[ 
- m p(m, \tau) 
+ (m + 1) p(m + 1, \tau) 
+ \lambda \sum_{m' = 0}^m G_{m-m'} p(m', \tau) - \lambda p(m, \tau)
\right],
\end{equation}
where we use $\sum_m \equiv \sum_{m=0}^\infty$. We can distribute the sum and
use the definition of $F(z, t)$ to obtain
\begin{equation}
{d F(z, \tau) \over d\tau} =
- \sum_m z^m m p(m, \tau)
+ \sum_m z^m (m + 1) p(m + 1, \tau)
+ \lambda \sum_m z^m \sum_{m'=0}^m G_{m-m'} p(m', \tau)
- \lambda F(z, \tau).
\label{eq:si_generating_01}
\end{equation}
We can make use of properties of the generating function to write everything in
terms of $F(z, \tau)$: the first term on the right hand side of
Eq.~\ref{eq:si_generating_01} can be rewritten as
\begin{align}
\sum_{m} z^{m} \cdot m \cdot p(m, \tau) &=
\sum_{m} z \frac{\partial z^{m}}{\partial z} p(m, \tau), \\
&=\sum_{m} z \frac{\partial}{\partial z}\left(z^{m} p(m, \tau)\right), \\
&=z \frac{\partial}{\partial z}\left(\sum_{m} z^{m} p(m, \tau)\right), \\
&= z {\partial F(z, \tau) \over \partial z}.
\end{align}
For the second term on the right hand side of Eq.~\ref{eq:si_generating_01} we
define $k \equiv m + 1$. This allows us to write
\begin{align}
\sum_{m=0}^{\infty} z^{m} \cdot(m+1) \cdot p(m+1, \tau) &=
\sum_{k=1}^{\infty} z^{k-1} \cdot k \cdot p(k, \tau), \\
&=z^{-1} \sum_{k=1}^{\infty} z^{k} \cdot k \cdot p(k, \tau), \\
&=z^{-1} \sum_{k=0}^{\infty} z^{k} \cdot k \cdot p(k, \tau), \\
&=z^{-1} \left(z \frac{\partial F(z)}{\partial z}\right), \\
&=\frac{\partial F(z)}{\partial z}.
\end{align}

@@

The third term in Eq.~\ref{eq:si_generating_01} is the most trouble. The trick
is to reverse the default order of the sums as
\begin{equation}
\sum_{m=0}^{\infty} \sum_{m'=0}^{m} = \sum_{m'=0}^{\infty} \sum_{m=m'}^{\infty}.
\end{equation}
To see the logic of the sum we point the reader to
Figure~\ref{figS2:sum_reindex}. The key is to notice that the double sum
$\sum_{m=0}^\infty \sum_{m'=0}^m$ is adding all possible pairs $(m, m')$ in the
lower triangle, so we can add the terms vertically as the original sum
indexing suggests, i.e.
\begin{equation}
\sum_{m=0}^{\infty} \sum_{m'=0}^{m} x_{(m, m')}= 
x_{(0, 0)} + x_{(1, 0)} + x_{(1, 1)} + x_{(2, 0)} + x_{(2, 1)} + x_{(2, 2)} + 
\ldots,
\end{equation}
where the variable $x$ is just a placeholder to indicate the order in which the
sum is taking place. But we can also add the terms horizontally as
\begin{equation}
\sum_{m'=0}^{\infty} \sum_{m=m'}^{\infty} x_{(m, m')} =
x_{(0, 0)} + x_{(1, 0)} + x_{(2, 0)} + \ldots + x_{(1,1)} + x_{(2, 1)} + \ldots,
\end{equation}
which still adds all of the lower triangle terms. Applying this reindexing
results in
\begin{align}
\lambda \sum_m z^m \sum_{m'=0}^m G_{m-m'} p(m', \tau) =
\lambda \sum_{m'=0}^{\infty} \sum_{m=m'}^{\infty} z^m 
\theta (1 - \theta)^{m-m'} p(m', \tau),
\end{align}
where we also substituted the definition of the geometric distribution $G_{k} =
\theta (1 - \theta)^k$. Redistributing the sums we can write
\begin{align}
\lambda \sum_{m'=0}^{\infty} \sum_{m=m'}^{\infty} z^m 
\theta (1 - \theta)^{m-m'} p(m', \tau) = 
\lambda \theta \sum_{n=0}^{\infty}(1-\theta)^{m'} P(m', \tau) 
\sum_{m=m'}^{\infty} \left[z (1-\theta)\right]^{m}.
\label{eq:si_generating_02}
\end{align}

The next step requires us to look slightly ahead into what we expect to obtain.
We are working on deriving an equation for the generating function $F(z, \tau)$
that when solved will allow us to compute what we care about, i.e. the
probability function $p(m, \tau)$. Upon finding the function for $F(z, \tau)$,
we will recover this probability distribution by evaluating derivatives
of $F(z, \tau)$ at $z=0$, whereas we can evaluate derivatives of
$F(z, \tau)$ at $z=1$ to instead recover the moments of the
distribution. The point here is that when the dust settles we
will evaluate $z$ to be less than or equal to one.
Furthermore, we know that the parameter of the geometric distribution $\theta$
must be strictly between zero and one. With these two facts we can safely state that
$| z (1 - \theta) | < 1$. Defining $n \equiv m - m'$ we rewrite the last sum in
Eq.~\ref{eq:si_generating_02} as
\begin{align}
\sum_{m=m'}^{\infty} \left[z (1-\theta)\right]^{m} &= 
\sum_{n=0}^{\infty} \left[z (1-\theta)\right]^{n + m'} \\
&= \left[ z (1 - \theta) \right]^{m'} 
\sum_{n=0}^{\infty} \left[ z (1 - \theta) \right]^{n} \\
&= \left[ z (1 - \theta) \right]^{m'} 
\left( {1 \over 1 - z (1 - \theta)} \right),
\end{align}
where we use the geometric series since, as stated before, $| z (1 - \theta) | <
1$. Putting these results together, the PDE for the generating function is
\begin{equation}
{\partial F \over \partial \tau} = 
{\partial F \over\partial z}
- z {\partial F \over \partial z} - \lambda F
+ \frac{\lambda\theta F}{1-z(1-\theta)}.
\end{equation}
Changing variables to $\xi=1-\theta$ and simplifying gives
\begin{align}
{\partial F \over \partial \tau} + (z - 1) {\partial F \over \partial z} = 
\frac{(z-1)\xi}{1-z\xi}\lambda F.
\label{eq:1state_unreg_015}
\end{align}

@@

To get at the mRNA distribution at steady state we first must solve
Eq.~\ref{eq:1state_unreg_015} setting the time derivative to zero. At
steady-state, the PDE reduces to the ODE
\begin{align}
\deriv[F]{z} = \frac{\xi}{1-z\xi}\lambda F,
\end{align}
which we can integrate as
\begin{align}
\int \frac{dF}{F} = \int \frac{\lambda\xi dz}{1-\xi z}.
\end{align}
The initial conditions for generating functions can be subtle and confusing. The
key fact follows from the definition
$F(z,t) = \sum_m z^m p(m,t)$.
Clearly normalization of the distribution requires that
$F(z=1, t) = \sum_m p(m,t) = 1$.
A subtlety is that sometimes the generating function may be undefined
\textit{at} $z=1$, in which case the limit as $z$ approaches $1$ from
below suffices to define the normalization condition.
We also warn the reader that, while it is frequently convenient to change
variables from $z$ to a different independent variable, one must
carefully track how the normalization condition transforms.

Continuing on, we evaluate the integrals (producing a constant $c$) which gives
\begin{align}
\ln F &= -\lambda \ln(1-\xi z) + c
\\
F &= \frac{c}{(1-\xi z)^\lambda}.
\end{align}
Only one choice for $c$ can satisfy initial conditions, producing
\begin{align}
F(z) = \left(\frac{1-\xi}{1-\xi z}\right)^\lambda
        = \left(\frac{\theta}{1 - z(1-\theta)}\right)^\lambda,
\label{eq:gen_fn}
\end{align}

@@

Let us move from a one-state promoter to a two-state promoter, where one state
has repressor bound and the other produces transcriptional bursts as above.
A schematic of this model is shown as model 5 in
Figure~\ref{fig1:means_cartoons}(C). Although now we have an equation for each
promoter state, otherwise the master equation reads similarly to the one-state
case, except with additional terms corresponding to transitions between promoter
states, namely
\begin{align}
{d\over dt}p_R(m,t) =& 
k_R^+ p_A(m,t) - k_R^- p_R(m,t)
        + (m+1)\gamma p_R(m+1,t) - m\gamma p_R(m,t)
\\
\begin{split}
{d\over dt}p_A(m,t) =& - k_R^+ p_A(m,t) + k_R^- p_R(m,t)
        + (m+1)\gamma p_A(m+1,t) - m\gamma p_A(m,t) 
\\
&- k_i p_A(m,t) + k_i \sum_{m^\prime=0}^m \theta(1-\theta)^{m-m^\prime} p_A(m^\prime,t),
\end{split}
\end{align}
where $p_R(m,t)$ is the probability of the system having $m$ mRNA copies and
having repressor bound to the promoter at time $t$, and $p_A$ is an analogous
probability to find the promoter without repressor bound. $k_R+$ and $k_R^-$
are, respectively, the rates at which repressors bind and unbind to and from the
promoter, and $\gamma$ is the mRNA degradation rate. $k_i$ is the rate at which
bursts initiate, and as before, the geometric distribution of burst sizes has
mean $b=(1-\theta)/\theta$.

@@

To verify this intuitively conjectured mapping between our problem and the
solution in~\cite{Shahrezaei2008}, we continue with a careful solution for the
mRNA distribution using probability generating functions, following the ideas
sketched in~\cite{Shahrezaei2008}. It is natural to nondimensionalize rates in
the problem by $\gamma$, or equivalently, this amounts to measuring time in
units of $\gamma^{-1}$. We are also only interested in steady state, so we set
the time derivatives to zero, giving
\begin{align}
0 =& k_R^+ p_A(m) - k_R^- p_R(m) + (m+1) p_R(m+1) - m p_R(m)
\\
\begin{split}
0 =& - k_R^+ p_A(m) + k_R^- p_R(m) + (m+1) p_A(m+1) - m p_A(m) 
\\
&- k_i p_A(m) + k_i \sum_{m^\prime=0}^m \theta(1-\theta)^{m-m^\prime} p_A(m^\prime),
\end{split}
\end{align}
where for convenience we kept the same notation for all rates, but these are
now expressed in units of mean mRNA lifetime $\gamma^{-1}$.

The probability generating function is defined as before in the constitutive
case, except now we must introduce a generating function for each promoter
state,
\begin{align}
f_A(z) = \sum_{m=0}^\infty z^m p_A(m),
\;
f_R(z) = \sum_{m=0}^\infty z^m p_R(m).
\end{align}
Our real objective is the generating function $f(z)$ that generates the mRNA
distribution $p(m)$, independent of what state the promoter is in. But since
$p(m) = p_A(m) + p_R(m)$, it follows too that $f(z) = f_A(z) + f_R(z)$.

As before we multiply both equations by $z^m$ and sum over all $m$. Each
individual term transforms exactly as did an analogous term in the constitutive
case, so the coupled ODEs for the generating functions read
\begin{align}
0 =& k_R^+ f_A(z) - k_R^- f_R(z) + \pderiv{z} f_R(z) - z \pderiv{z} f_R(z)
\\
\begin{split}
0 =&  - k_R^+ f_A(z) + k_R^- f_R(z) + \pderiv{z} f_A(z) - z \pderiv{z} f_A(z)
\\
&- k_i f_A(z) + k_i \frac{\theta}{1-z(1-\theta)} f_A(z),
\end{split}
\end{align}
and after changing variables $\xi = 1 - \theta$ as before and rearranging, we
have
\begin{align}
0 &= k_R^+ f_A(z) - k_R^- f_R(z) + (1-z) \pderiv{z} f_R(z)
\\
0 &=  - k_R^+ f_A(z) + k_R^- f_R(z) + (1 - z) \pderiv{z} f_A(z)
+ k_i \frac{(z-1)\xi}{1-z\xi} f_A(z),
\end{align}
We can transform this problem from two coupled first-order ODEs to a single
second-order ODE by solving for $f_A$ in the first and plugging into the second,
giving
\begin{align}
\begin{split}
0 = (1&-z) \pderiv[f_R]{z}
+ \frac{1-z}{k_R^+}
        \left(k_R^- \pderiv[f_R]{z} + \pderiv[f_R]{z} +(z-1) \psecderiv[f_R]{z}\right)
\\
&+ \frac{k_i}{k_R^+} \frac{(z-1)\xi}{1-z\xi}
        \left(k_R^- f_R + (z-1) \pderiv[f_R]{z}\right),
\end{split}
\end{align}
where, to reduce notational clutter, we have dropped the explicit $z$ dependence
of $f_A$ and $f_R$. Simplifying we have
\begin{align}
0 = \psecderiv[f_R]{z}
        - \left(\frac{k_i\xi}{1-z\xi}
                + \frac{1 + k_R^- + k_R^+}{1-z}
        \right)\pderiv[f_R]{z}
        + \frac{k_i k_R^- \xi}{(1-z\xi)(1-z)}f_R.
\end{align}
This can be recognized as the hypergeometric differential equation, with
singularities at $z=1$, $z=\xi^{-1}$, and $z=\infty$. The latter can be verified
by a change of variables from $z$ to $x=1/z$, being careful with the chain rule,
and noting that $z=\infty$ is a singular point if and only if $x=1/z=0$ is a
singular point.

@@

\begin{align}

@@

\end{align}

@@

${_2F_1}$ and differ only in the sign preceeding the radical.

>>>>

To clean up Equation~\ref{eq:si_master_deltat} we can send the first term on the
right hand side to the left, and divide both sides by $\Delta t$. Upon taking
the limit where $\Delta t \rightarrow 0$ we can write
\begin{equation}
{d \over dt}p(m, t) = (m + 1) \gamma p(m + 1, t)
- m \gamma p(m, t)
+ k_i \sum_{m'=0}^{m-1} G_{m-m'} p(m', t) 
- k_i \sum_{m'=m + 1}^{\infty} G_{m'-m} p(m, t).
\end{equation}
Furthermore, given that the timescale for this equation is set by the mRNA
degradation rate $\gamma$ we can divide both sides by this rate, obtaining
\begin{equation}
{d \over d\tau}p(m, \tau) = (m + 1) p(m + 1, \tau)
- m p(m, \tau)
+ \lambda \sum_{m'=0}^{m-1} G_{m-m'} p(m', \tau) 
- \lambda \sum_{m'=m + 1}^{\infty} G_{m'-m} p(m, \tau),
\label{eq:si_master_ode}
\end{equation}
where we defined $\tau \equiv t \times \gamma$, and $\lambda \equiv k_i/\gamma$.
The last term in Eq.~\ref{eq:si_master_ode} sums all burst sizes except for a 
burst of size zero. We can re-index the sum to include this term, obtaining
\begin{equation}
\lambda \sum_{m'=m + 1}^{\infty} G_{m'-m} p(m, \tau) = \lambda p(m, t) \left[
\underbrace{\sum_{m'={m}}^{\infty}G_{m'-m}}
_{\text{re-index sum to include burst size zero}} -
\underbrace{G_0}_{\text{subtract extra added term}}\right].
\end{equation}
Given the normalization constraint of the geometric distribution, adding the
probability of all possible burst sizes --including size zero since we
re-indexed the sum-- allows us to write
\begin{equation}
\sum_{m'=m}^{\infty}G_{m'-m} - G_0 = 1 - G_0.
\end{equation}
Substituting this into Eq.~\ref{eq:si_master_ode} results in
\begin{equation}
{d \over d\tau}p(m, \tau) = (m + 1) p(m + 1, \tau)
- m p(m, \tau)
+ \lambda \sum_{m'=0}^{m-1} G_{m-m'} p(m', \tau) 
- \lambda p(m, \tau) \left[ 1 - G_0 \right].
\label{eq:si_master_ode_2}
\end{equation}
To finally get at a more compact version of the equation notice that the third
term in Eq.~\ref{eq:si_master_ode_2} includes burst from size $m'-m = 1$ to size
$m' - m = m$. We can include the term $p(m, t) G_0$ in the sum which allows
bursts of size $m' - m = 0$. This results in our final form for the chemical
master equation
\begin{equation}
{d \over d\tau}p(m, \tau) = 
(m + 1) p(m+1, \tau)
- m p(m, \tau) - 
\lambda p(m, \tau)
+ \lambda \sum_{m^\prime=0}^m G_{m-m^\prime} p(m^\prime, \tau).
\label{eq:si_master_unreg}
\end{equation}

In order to solve Eq.~\ref{eq:si_master_unreg} we will use the generating
function method~\cite{vanKampen2007}. The probability generating function
is defined as
\begin{equation}
F(z,t) = \sum_{m=0}^\infty z^m p(m,t),
\end{equation}
where $z$ is just a dummy variable that will help us later on to obtain the
moments of the distribution. Let us now multiply both sides of
Eq.~\ref{eq:si_master_unreg} by $z^m$ and sum over all $m$
\begin{equation}
\sum_m z^m {d \over d\tau} p(m, \tau) = 
\sum_m z^m \left[ 
- m p(m, \tau) 
+ (m + 1) p(m + 1, \tau) 
+ \lambda \sum_{m' = 0}^m G_{m-m'} p(m', \tau) - \lambda p(m, \tau)
\right],
\end{equation}
where we use $\sum_m \equiv \sum_{m=0}^\infty$. We can distribute the sum and
use the definition of $F(z, t)$ to obtain
\begin{equation}
{d F(z, \tau) \over d\tau} =
- \sum_m z^m m p(m, \tau)
+ \sum_m z^m (m + 1) p(m + 1, \tau)
+ \lambda \sum_m z^m \sum_{m'=0}^m G_{m-m'} p(m', \tau)
- \lambda F(z, \tau).
\label{eq:si_generating_01}
\end{equation}
We can make use of properties of the generating function to write everything in
terms of $F(z, \tau)$: the first term on the right hand side of
Eq.~\ref{eq:si_generating_01} can be rewritten as
\begin{eqnarray}
\sum_{m} z^{m} \cdot m \cdot p(m, \tau) &=&
\sum_{m} z \frac{\partial z^{m}}{\partial z} p(m, \tau), \\
&=&\sum_{m} z \frac{\partial}{\partial z}\left(z^{m} p(m, \tau)\right), \\
&=&z \frac{\partial}{\partial z}\left(\sum_{m} z^{m} p(m, \tau)\right), \\
&=& z {\partial F(z, \tau) \over \partial z}.
\end{eqnarray}
For the second term on the right hand side of Eq.~\ref{eq:si_generating_01} we
define $k \equiv m + 1$. This allows us to write
\begin{eqnarray}
\sum_{m=0}^{\infty} z^{m} \cdot(m+1) \cdot p(m+1, \tau) &=&
\sum_{k=1}^{\infty} z^{k-1} \cdot k \cdot p(k, \tau), \\
&=&z^{-1} \sum_{k=1}^{\infty} z^{k} \cdot k \cdot p(k, \tau), \\
&=&z^{-1} \sum_{k=0}^{\infty} z^{k} \cdot k \cdot p(k, \tau), \\
&=&z^{-1} \left(z \frac{\partial F(z)}{\partial z}\right), \\
&=&\frac{\partial F(z)}{\partial z}.
\end{eqnarray}

@@

The third term in Eq.~\ref{eq:si_generating_01} is the most trouble. The trick
is to reverse the default order of the sums as
\begin{equation}
\sum_{m=0}^{\infty} \sum_{m'=0}^{m} = \sum_{m'=0}^{\infty} \sum_{m=m'}^{\infty}.
\end{equation}
To see the logic of the sum we point the reader to
Figure~\ref{figS2:sum_reindex}. The key is to notice that the double sum
$\sum_{m=0}^\infty \sum_{m'=0}^m$ is adding all possible pairs $(m, m')$ in the
lower triangle, so we can add the terms vertically as the original sum
indexing suggests, i.e.
\begin{equation}
\sum_{m=0}^{\infty} \sum_{m'=0}^{m} x_{(m, m')}= 
x_{(0, 0)} + x_{(1, 0)} + x_{(1, 1)} + x_{(2, 0)} + x_{(2, 1)} + x_{(2, 2)} + 
\ldots,
\end{equation}
where the variable $x$ is just a placeholder to indicate the order in which the
sum is taking place. But we can also add the terms horizontally as
\begin{equation}
\sum_{m'=0}^{\infty} \sum_{m=m'}^{\infty} x_{(m, m')} =
x_{(0, 0)} + x_{(1, 0)} + x_{(2, 0)} + \ldots + x_{(1,1)} + x_{(2, 1)} + \ldots,
\end{equation}
which still adds all of the lower triangle terms. Applying this reindexing
results in
\begin{equation}
\lambda \sum_m z^m \sum_{m'=0}^m G_{m-m'} p(m', \tau) =
\lambda \sum_{m'=0}^{\infty} \sum_{m=m'}^{\infty} z^m 
\theta (1 - \theta)^{m-m'} p(m', \tau),
\end{equation}
where we also substituted the definition of the geometric distribution $G_{k} =
\theta (1 - \theta)^k$. Redistributing the sums we can write
\begin{equation}
\lambda \sum_{m'=0}^{\infty} \sum_{m=m'}^{\infty} z^m 
\theta (1 - \theta)^{m-m'} p(m', \tau) = 
\lambda \theta \sum_{n=0}^{\infty}(1-\theta)^{m'} P(m', \tau) 
\sum_{m=m'}^{\infty} \left[z (1-\theta)\right]^{m}.
\label{eq:si_generating_02}
\end{equation}

The next step requires us to look slightly ahead into what we expect to obtain.
We are working on deriving an equation for the generating function $F(z, \tau)$
that when solved will allow us to compute what we care about, i.e. the
probability function $p(m, \tau)$. Upon finding the function for $F(z, \tau)$,
we will recover this probability distribution by evaluating derivatives
of $F(z, \tau)$ at $z=0$, whereas we can evaluate derivatives of
$F(z, \tau)$ at $z=1$ to instead recover the moments of the
distribution. The point here is that when the dust settles we
will evaluate $z$ to be less than or equal to one.
Furthermore, we know that the parameter of the geometric distribution $\theta$
must be strictly between zero and one. With these two facts we can safely state that
$| z (1 - \theta) | < 1$. Defining $n \equiv m - m'$ we rewrite the last sum in
Eq.~\ref{eq:si_generating_02} as
\begin{eqnarray}
\sum_{m=m'}^{\infty} \left[z (1-\theta)\right]^{m} &=&
\sum_{n=0}^{\infty} \left[z (1-\theta)\right]^{n + m'} \\
&=& \left[ z (1 - \theta) \right]^{m'} 
\sum_{n=0}^{\infty} \left[ z (1 - \theta) \right]^{n} \\
&=& \left[ z (1 - \theta) \right]^{m'} 
\left( {1 \over 1 - z (1 - \theta)} \right),
\end{eqnarray}
where we use the geometric series since, as stated before, $| z (1 - \theta) | <
1$. Putting these results together, the PDE for the generating function is
\begin{equation}
{\partial F \over \partial \tau} = 
{\partial F \over\partial z}
- z {\partial F \over \partial z} - \lambda F
+ \frac{\lambda\theta F}{1-z(1-\theta)}.
\end{equation}
Changing variables to $\xi=1-\theta$ and simplifying gives
\begin{align}
{\partial F \over \partial \tau} + (z - 1) {\partial F \over \partial z} = 
\frac{(z-1)\xi}{1-z\xi}\lambda F.
\label{eq:1state_unreg_015}
\end{align}

@@

To get at the mRNA distribution at steady state we first must solve
Eq.~\ref{eq:1state_unreg_015} setting the time derivative to zero. At
steady-state, the PDE reduces to the ODE
\begin{equation}
\deriv[F]{z} = \frac{\xi}{1-z\xi}\lambda F,
\end{equation}
which we can integrate as
\begin{equation}
\int \frac{dF}{F} = \int \frac{\lambda\xi dz}{1-\xi z}.
\end{equation}
The initial conditions for generating functions can be subtle and confusing. The
key fact follows from the definition
$F(z,t) = \sum_m z^m p(m,t)$.
Clearly normalization of the distribution requires that
$F(z=1, t) = \sum_m p(m,t) = 1$.
A subtlety is that sometimes the generating function may be undefined
\textit{at} $z=1$, in which case the limit as $z$ approaches $1$ from
below suffices to define the normalization condition.
We also warn the reader that, while it is frequently convenient to change
variables from $z$ to a different independent variable, one must
carefully track how the normalization condition transforms.

Continuing on, we evaluate the integrals (producing a constant $c$) which gives
\begin{eqnarray}
\ln F &=& -\lambda \ln(1-\xi z) + c
\\
F &=& \frac{c}{(1-\xi z)^\lambda}.
\end{eqnarray}
Only one choice for $c$ can satisfy initial conditions, producing
\begin{equation}
F(z) = \left(\frac{1-\xi}{1-\xi z}\right)^\lambda
        = \left(\frac{\theta}{1 - z(1-\theta)}\right)^\lambda,
\label{eq:gen_fn}
\end{equation}

@@

Let us move from a one-state promoter to a two-state promoter, where one state
has repressor bound and the other produces transcriptional bursts as above.
A schematic of this model is shown as model 5 in
Figure~\ref{fig1:means_cartoons}(C). Although now we have an equation for each
promoter state, otherwise the master equation reads similarly to the one-state
case, except with additional terms corresponding to transitions between promoter
states, namely
\begin{equation}
{d\over dt}p_R(m,t) =
k_R^+ p_A(m,t) - k_R^- p_R(m,t)
        + (m+1)\gamma p_R(m+1,t) - m\gamma p_R(m,t),
\end{equation}
and
\begin{equation}
\begin{split}
{d\over dt}p_A(m,t) = - k_R^+ p_A(m,t) + k_R^- p_R(m,t)
        + (m+1)\gamma p_A(m+1,t) - m\gamma p_A(m,t) 
\\
- k_i p_A(m,t) + k_i \sum_{m^\prime=0}^m \theta(1-\theta)^{m-m^\prime} p_A(m^\prime,t),
\end{split}
\end{equation}
where $p_R(m,t)$ is the probability of the system having $m$ mRNA copies and
having repressor bound to the promoter at time $t$, and $p_A$ is an analogous
probability to find the promoter without repressor bound. $k_R+$ and $k_R^-$
are, respectively, the rates at which repressors bind and unbind to and from the
promoter, and $\gamma$ is the mRNA degradation rate. $k_i$ is the rate at which
bursts initiate, and as before, the geometric distribution of burst sizes has
mean $b=(1-\theta)/\theta$.

@@

To verify this intuitively conjectured mapping between our problem and the
solution in~\cite{Shahrezaei2008}, we continue with a careful solution for the
mRNA distribution using probability generating functions, following the ideas
sketched in~\cite{Shahrezaei2008}. It is natural to nondimensionalize rates in
the problem by $\gamma$, or equivalently, this amounts to measuring time in
units of $\gamma^{-1}$. We are also only interested in steady state, so we set
the time derivatives to zero, giving
\begin{equation}
0 = k_R^+ p_A(m) - k_R^- p_R(m) + (m+1) p_R(m+1) - m p_R(m),
\end{equation}
and
\begin{equation}
\begin{split}
0 = - k_R^+ p_A(m) + k_R^- p_R(m) + (m+1) p_A(m+1) - m p_A(m) 
\\
- k_i p_A(m) + k_i \sum_{m^\prime=0}^m \theta(1-\theta)^{m-m^\prime} p_A(m^\prime),
\end{split}
\end{equation}
where for convenience we kept the same notation for all rates, but these are
now expressed in units of mean mRNA lifetime $\gamma^{-1}$.

The probability generating function is defined as before in the constitutive
case, except now we must introduce a generating function for each promoter
state,
\begin{equation}
f_A(z) = \sum_{m=0}^\infty z^m p_A(m),
\;
f_R(z) = \sum_{m=0}^\infty z^m p_R(m).
\end{equation}
Our real objective is the generating function $f(z)$ that generates the mRNA
distribution $p(m)$, independent of what state the promoter is in. But since
$p(m) = p_A(m) + p_R(m)$, it follows too that $f(z) = f_A(z) + f_R(z)$.

As before we multiply both equations by $z^m$ and sum over all $m$. Each
individual term transforms exactly as did an analogous term in the constitutive
case, so the coupled ODEs for the generating functions read
\begin{equation}
0 = k_R^+ f_A(z) - k_R^- f_R(z) + \pderiv{z} f_R(z) - z \pderiv{z} f_R(z),
\end{equation}
and
\begin{equation}
\begin{split}
0 =&  - k_R^+ f_A(z) + k_R^- f_R(z) + \pderiv{z} f_A(z) - z \pderiv{z} f_A(z)
\\
&- k_i f_A(z) + k_i \frac{\theta}{1-z(1-\theta)} f_A(z),
\end{split}
\end{equation}
and after changing variables $\xi = 1 - \theta$ as before and rearranging, we
have
\begin{eqnarray}
0 &=& k_R^+ f_A(z) - k_R^- f_R(z) + (1-z) \pderiv{z} f_R(z)
\\
0 &=&  - k_R^+ f_A(z) + k_R^- f_R(z) + (1 - z) \pderiv{z} f_A(z)
+ k_i \frac{(z-1)\xi}{1-z\xi} f_A(z),
\end{eqnarray}
We can transform this problem from two coupled first-order ODEs to a single
second-order ODE by solving for $f_A$ in the first and plugging into the second,
giving
\begin{equation}
\begin{split}
0 = (1&-z) \pderiv[f_R]{z}
+ \frac{1-z}{k_R^+}
        \left(k_R^- \pderiv[f_R]{z} + \pderiv[f_R]{z} +(z-1) \psecderiv[f_R]{z}\right)
\\
&+ \frac{k_i}{k_R^+} \frac{(z-1)\xi}{1-z\xi}
        \left(k_R^- f_R + (z-1) \pderiv[f_R]{z}\right),
\end{split}
\end{equation}
where, to reduce notational clutter, we have dropped the explicit $z$ dependence
of $f_A$ and $f_R$. Simplifying we have
\begin{equation}
0 = \psecderiv[f_R]{z}
        - \left(\frac{k_i\xi}{1-z\xi}
                + \frac{1 + k_R^- + k_R^+}{1-z}
        \right)\pderiv[f_R]{z}
        + \frac{k_i k_R^- \xi}{(1-z\xi)(1-z)}f_R.
\end{equation}
This can be recognized as the hypergeometric differential equation, with
singularities at $z=1$, $z=\xi^{-1}$, and $z=\infty$. The latter can be verified
by a change of variables from $z$ to $x=1/z$, being careful with the chain rule,
and noting that $z=\infty$ is a singular point if and only if $x=1/z=0$ is a
singular point.

@@

\begin{equation}

@@

\end{equation}

@@

${_2F_1}$ and differ only in the sign preceding the radical.
